Clustering in a hyperbolic model of complex networks.

### 2020-03-10 paper.tex
Edits by Tobias

%\usepackage{showkeys}
\usepackage{tocloft}

\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{bm}
\usepackage[mathscr]{eucal}
\usepackage{a4wide}
\usepackage{authblk}
\usepackage{appendix}
\usepackage{hyperref}
\hypersetup{
    colorlinks = true,
    urlcolor = {magenta},
    citecolor = {green},
    linkcolor = {blue}
}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,backgrounds,patterns,decorations.markings,decorations.pathreplacing,intersections}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}

@@

\newcommand{\BallHyp}[1]{\Bcal\left(#1\right)}		%	The image of the hyperbolic ball around \Psi^{-1}(p)
\newcommand{\BallPo}[1]{\Bcal_{\infty}\left(#1\right)}		%	The neighborhood of a point p in the infinite limit model
\newcommand{\BallPon}[1]{\Bcal_{\text{box}}\left(#1\right)}	%	The neighborhood of a point p in the finit box model
\newcommand{\FatBallHyp}[1]{\check{\Bcal}_{\Po}(#1)}		%
\newcommand{\BallSym}[1]{\Bcal_{\Po \bigtriangleup \infty}\left(#1\right)}	%	Symmetric difference between balls in the 
																			%	hyperbolic and infinite limit model
\newcommand{\BallInter}[1]{\Bcal_{\Po \cap \text{box}}\left(#1\right)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Don't mess with the tobimacros !!!!!							  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\eR}[0]{\ensuremath{{\mathbb R}}}
\newcommand{\eN}[0]{\ensuremath{{\mathbb N}}}
\newcommand{\Ee}[0]{\ensuremath{{\mathbb E}}}
\newcommand{\Haa}{\ensuremath{{\mathbb H}}}
\newcommand{\GPo}[0]{\ensuremath{G_{\Po}}}
\newcommand{\Gbox}[0]{\ensuremath{G_{\mathrm{box}}}}
\newcommand{\GPoH}{\ensuremath{G_{Po,H}}}
\newcommand{\GboxH}{\ensuremath{G_{box,H}}}
\newcommand{\Ginf}[0]{\ensuremath{G_{\infty}}}
\newcommand{\isd}{\stackrel{\text{d}}{=}}
\newcommand{\Xcal}{\ensuremath{{\mathcal X}}}
\newcommand{\Ycal}{\ensuremath{{\mathcal Y}}}
\newcommand{\Fcal}{\ensuremath{\mathcal{F}}}
\newcommand{\Qcal}{\ensuremath{{\mathcal Q}}}
\newenvironment{proofof}[1]{\vspace{1ex}\noindent{\bf Proof of #1:}}{\hspace*{\fill}$\blacksquare$\vspace{1ex}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%								For leaving comments to each other:								  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Here we show that the clustering coefficient tends in probability to a constant $\gamma$ that we give explicitly as a closed form expression in terms of $\alpha, \nu$ and certain special functions. This improves earlier work by Gugelmann et al., who proved that the clustering coefficient remains bounded away from zero with high probability, but left open the issue of convergence to a limiting constant. Similarly, we are able to show that $c(k)$, the average clustering coefficient over all vertices of degree exactly $k$, tends in probability to a limit $\gamma(k)$ which we give explicitly as a closed form expression in terms of $\alpha, \nu$ and certain special functions. We are able to extend this last result also to sequences $(k_n)_n$ where $k_n$ grows as a function of $n$. Our results show that $\gamma(k)$ scales differently, as $k$ grows, for different ranges of $\alpha$. More precisely, there exists constants $Q_{\alpha,\nu}$ depending on $\alpha$ and $\nu$, such that as $k \to \infty$, $\gamma(k) \sim Q_{\alpha,\nu} k^{2 - 4\alpha}$ if $\frac{1}{2} < \alpha < \frac{3}{4}$, $\gamma(k) \sim Q_{\alpha,\nu} \log(k) k^{-1} $ if $\alpha=\frac{3}{4}$ and $\gamma = Q_{\alpha,\nu} k^{-1}$ when $\alpha > \frac{3}{4}$. These results contradict a prediction of Krioukov et al., which stated that the limiting values $\gamma(k)$ should always scale with $k^{-1}$ as we let $k$ grow. 
\end{abstract}

@@

\TM{ Maybe a ``conclusion and further work'' section here,  before the references and appendix? \\
Also, appendix usually comes after references in my experience.}

>>>>

%\usepackage{showkeys}
\usepackage{tocloft}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{bm}
\usepackage[mathscr]{eucal}
\usepackage{a4wide}
\usepackage{authblk}
\usepackage{appendix}
\usepackage{hyperref}
\hypersetup{
    colorlinks = true,
    urlcolor = {magenta},
    citecolor = {green},
    linkcolor = {blue}
}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,backgrounds,patterns,decorations.markings,decorations.pathreplacing,intersections}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}

@@

\newcommand{\BallHyp}[1]{\Bcal\left(#1\right)}		%	The image of the hyperbolic ball around \Psi^{-1}(p)
\newcommand{\BallPo}[1]{\Bcal_{\infty}\left(#1\right)}		%	The neighborhood of a point p in the infinite limit model
\newcommand{\BallPon}[1]{\Bcal_{\text{box}}\left(#1\right)}	%	The neighborhood of a point p in the finit box model
\newcommand{\FatBallHyp}[1]{\check{\Bcal}_{\Po}(#1)}		%
\newcommand{\BallSym}[1]{\Bcal_{\Po \bigtriangleup \infty}\left(#1\right)}	%	Symmetric difference between balls in the 
\newcommand{\BallH}[1]{\Bcal_{H}\left(#1\right)}	 % Ball up to height H.
%	hyperbolic and infinite limit model
\newcommand{\BallInter}[1]{\Bcal_{\Po \cap \text{box}}\left(#1\right)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Don't mess with the tobimacros !!!!!							  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\eR}[0]{\ensuremath{{\mathbb R}}}
\newcommand{\pmf}[0]{\ensuremath{{\pi}}}
\newcommand{\eN}[0]{\ensuremath{{\mathbb N}}}
\newcommand{\Ee}[0]{\ensuremath{{\mathbb E}}}
\newcommand{\Haa}{\ensuremath{{\mathbb H}}}
\newcommand{\GPo}[0]{\ensuremath{G_{\Po}}}
\newcommand{\Gbox}[0]{\ensuremath{G_{\mathrm{box}}}}
%\newcommand{\GboxH}[0]{\ensuremath{G_{\mathrm{box},\leq R/4}}}
\newcommand{\GboxH}[0]{\ensuremath{G_{\mathrm{box}-}}}
%\newcommand{\GboxH}[0]{\ensuremath{G_{-}}}
\newcommand{\Ginf}[0]{\ensuremath{G_{\infty}}}
\newcommand{\isd}{\stackrel{\text{d}}{=}}
\newcommand{\Xcal}{\ensuremath{{\mathcal X}}}
\newcommand{\Ycal}{\ensuremath{{\mathcal Y}}}
\newcommand{\Fcal}{\ensuremath{\mathcal{F}}}
\newcommand{\Qcal}{\ensuremath{{\mathcal Q}}}
\newenvironment{proofof}[1]{\vspace{1ex}\noindent{\bf Proof of #1.}}{\hspace*{\fill}$\blacksquare$\vspace{1ex}}
\renewenvironment{proof}{\vspace{1ex}\noindent{\bf Proof.}}{\hspace*{\fill}$\blacksquare$\vspace{1ex}}
\DeclareMathOperator{\Var}{Var}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%								For leaving comments to each other:								  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Here we show that the clustering coefficient tends in probability to a constant $\gamma$ that we give explicitly as a closed form 
expression in terms of $\alpha, \nu$ and certain special functions. 
This improves earlier work by Gugelmann et al., who proved that the clustering coefficient remains bounded away from zero 
with high probability, but left open the issue of convergence to a limiting constant. Similarly, we are able to show that 
$c(k)$, the average clustering coefficient over all vertices of degree exactly $k$, tends in probability to a limit 
$\gamma(k)$ which we give explicitly as a closed form expression in terms of $\alpha, \nu$ and certain special functions. 
We are able to extend this last result also to sequences $(k_n)_n$ where $k_n$ grows as a function of $n$. 
Our results show that $\gamma(k)$ scales differently, as $k$ grows, for different ranges of $\alpha$. 
More precisely, there exists constants $c_{\alpha,\nu}$ depending on $\alpha$ and $\nu$, such that as $k \to \infty$, 
$\gamma(k) \sim c_{\alpha,\nu} \cdot k^{2 - 4\alpha}$ if $\frac{1}{2} < \alpha < \frac{3}{4}$, 
$\gamma(k) \sim c_{\alpha,\nu} \cdot \log(k) \cdot k^{-1} $ 
if $\alpha=\frac{3}{4}$ and $\gamma(k) \sim c_{\alpha,\nu} \cdot k^{-1}$ when $\alpha > \frac{3}{4}$. 
These results contradict a conjecture of Krioukov et al., which stated that the limiting values $\gamma(k)$ should always scale 
with $k^{-1}$ as we let $k$ grow. 
\end{abstract}

@@

### 2020-03-10 preliminaries.tex
Edits by Tobias

For the definition of the finite graph, recall that in the definition of the KPKVB model we set $R = 2\log(n/\nu)$.
We consider the box $\Rcal = (-\frac{\pi}{2}e^{R/2}, \frac{\pi}{2}e^{R/2}] \times (0, R]$ in $\eR^2$. 
Then the \emph{finite box model} $\Gbox := \Gbox(n;\alpha, \nu)$ has vertex set $\Vbox := \mathcal{P} \cap \Rcal$ and edge set such that
\[
	pp' \in E(\Gbox(n;\alpha, \nu)) \iff |x - x'|_{\pi e^{R/2}} \leq e^{\frac{y + y'}{2}},
\]
where $|x|_{r} = \min( |x|, r - |x|)$ for $-r\leq x\leq r$. Using $|.|_{\pi e^{R/2}}$ instead of $|.|$ results in the left and right boundaries of the box $\Rcal$ getting identified, which in particular makes the model invariant under horizontal shifts and reflections in vertical lines. The graph $\Gbox$ can thus be seen as a subgraph of $\Ginf$ induced on $\Vbox$, with some additional edges caused by the identification of the boundaries.

@@

\begin{remark}[Notion convention for points]
We will often be working with the finite box graph $\GPo$ or the infinite graph $\Ginf$, whose nodes are points in $\R \times \R_+$. For any point $p \in \R \times \R_+$ we will always use $p = (x,y)$. When considering different points $p,p^\prime \in \R \times \R_+$, we will use primed coordinates to refer to $p^\prime$, i.e. $p^\prime = (x^\prime ,y^\prime)$, and similar with subscripts, i.e. $p_i = (x_i,y_i)$.
\end{remark}

@@

A very useful tool for analyzing subgraph counts, and their generalizations, in the 
setting of the Poissonized random geometric graphs, and in particular the Poissonized KPKVB model and the box model is the \emph{Campbell-Mecke formula}. We use a specific incarnation, which follows from the Palm theory of Poisson point processes on metric spaces, see~\cite{last2017lectures}. For this consider a Poisson point process $\Pcal$ on some metric space $\mathcal{M}$ with density $\mu$ and let $\mathcal{N}$ denote the set of all possible point configurations in $\mathcal{M}$, equipped with the sigma algebra of the process $\Pcal$. Then, for any natural number $k$ and measurable function $h : \R^k \times \mathcal{N} \to \R$,
\begin{equation}\label{eq:def_campbell_mecke}
	\Exp{\sum_{p_1, \dots, p_k \in \Pcal}^{\ne} h(p_1, \dots, p_k, \Pcal)}
	= \int_{\mathcal{M}} \dots \int_{\mathcal{M}} \Exp{h(x_1,\dots, x_k, \Pcal} \dd \mu(x_1) \dots \dd \mu(x_k),
\end{equation}
where the sum is over all distinct point $p_1, \dots, p_k \in \Pcal$.

@@

For our application these Chernoff bounds imply that if $y$ is such that $\hat{\mu}(y)$ is far from $k_n$ then $\Prob{\Po(\hat{\mu}(y)) = k_n}$ becomes very small. To be more specific, we define for any $k \ge 0$ and $C > 0$,
\begin{equation}\label{eq:def_y_k_C}
	y_{k,C}^\pm = 2 \log\left(\frac{k \pm C \sqrt{k \log(k)}}{\xi}\right),
\end{equation}
where we set $y_{k,C}^- = 0$ if $k - C \sqrt{k \log(k)} < \xi$. For convenience we write $\Kcal_C(k_n) := [y_{k_n,C}^-, y_{k_n,C}^+]$. Then we can show that for all $y$ outside $\Kcal_C(k_n)$
\begin{equation}\label{eq:chernoff_bound_degrees}
	\Prob{\Po(\hat{\mu}(y)) = k_n} \le \bigO{k_n^{-\frac{C^2}{2}}}.
\end{equation}
Since we can select $C$ to be as big as we want we can make this error as small as needed. This implies that then the main contribution to the integral~\eqref{eq:example_poisson_integral} comes from those `'heights" $y$ that are in the interval $\Kcal_C(k_n)$. In other words, the main contribution is concentrated around the heights $y$ for which $\mu(y) = k_n$. We thus refer to this as the concentration of heights result. More precisely, we prove the following.

>>>>

Recall that in the definition of the KPKVB model we set $R = 2\log(n/\nu)$.
We consider the box $\Rcal = (-\frac{\pi}{2}e^{R/2}, \frac{\pi}{2}e^{R/2}] \times (0, R]$ in $\eR^2$. 
Then the \emph{finite box model} $\Gbox = \Gbox(n;\alpha, \nu)$ has vertex set $\Vbox := \mathcal{P} \cap \Rcal$ and edge set such that
\[
	pp' \in E(\Gbox) \iff |x - x'|_{\pi e^{R/2}} \leq e^{\frac{y + y'}{2}},
\]
where $|x|_{r} = \min( |x|, r - |x|)$ for $-r\leq x\leq r$. Using $|.|_{\pi e^{R/2}}$ instead of $|.|$ results in the left and right boundaries of the box $\Rcal$ getting identified, which in particular makes the model invariant under horizontal shifts and reflections in vertical lines. 
The graph $\Gbox$ can thus be seen as a subgraph of $\Ginf$ induced on $\Vbox$, with some additional edges caused by the identification of the boundaries.

@@

\begin{remark}[Notational convention for points]
We will often be working with the finite box graph $\GPo$ or the infinite graph $\Ginf$, whose nodes are points 
in $\R \times \R_+$. For any point $p \in \R \times \R_+$ we will always use $p = (x,y)$. 
When considering different points $p,p^\prime \in \R \times \R_+$, we will use primed coordinates to refer 
to $p^\prime$, i.e. $p^\prime = (x^\prime ,y^\prime)$, and similar with subscripts, i.e. $p_i = (x_i,y_i)$.
\end{remark}

@@

A useful tool for analyzing subgraph counts, and their generalizations, in the 
setting of Poissonized random geometric graphs, and in particular the Poissonized KPKVB model and the box 
model is the \emph{Campbell-Mecke formula}. 
We use a specific incarnation, which follows from the Palm theory of Poisson point processes on metric 
spaces, see~\cite{last2017lectures}. For this consider a Poisson point process $\Pcal$ on some metric 
space $\mathcal{M}$ with density $\mu$ and let $\mathcal{N}$ denote the set of all possible point configurations 
in $\mathcal{M}$, equipped with the sigma algebra of the process $\Pcal$. 
Then, for any natural number $k$ and measurable function $h : \R^k \times \mathcal{N} \to \R$,

\begin{equation}\label{eq:def_campbell_mecke}
\begin{array}{c}
\displaystyle	\Exp{\sum_{p_1, \dots, p_k \in \Pcal, \atop \text{distinct}} h(p_1, \dots, p_k, \Pcal)} \\
= \\
\displaystyle \int_{\mathcal{M}} \dots \int_{\mathcal{M}} \Exp{h(x_1,\dots, x_k, \Pcal\cup\{x_1,\dots,x_k\})} \mu(dx_1)\dots\mu(dx_k).
\end{array}
\end{equation}

@@

For our application these Chernoff bounds imply that if $y$ is such that $\hat{\mu}(y)$ is far from $k_n$ then $\Prob{\Po(\hat{\mu}(y)) = k_n}$ becomes very small. To be more specific, we define for any $k \ge 0$ and $C > 0$,
\begin{equation}\label{eq:def_y_k_C}
	y_{k,C}^\pm = 2 \log\left(\frac{k \pm C \sqrt{k \log(k)}}{\xi}\right),
\end{equation}
where we set $y_{k,C}^- = 0$ if $k - C \sqrt{k \log(k)} < \xi$ and likewise if $k + C \sqrt{k \log(k)} < \xi$ we set $y_{k,C}^+=0$, but  note that as we consider $k\rightarrow\infty$, we can assume that this case does not occur. For convenience we write $\Kcal_C(k_n) := [y_{k_n,C}^-, y_{k_n,C}^+]$. Then we can show that for all $y$ outside $\Kcal_C(k_n)$
\begin{equation}\label{eq:chernoff_bound_degrees}
	\Prob{\Po(\hat{\mu}(y)) = k_n} \le \bigO{k_n^{-\frac{C^2}{2}}}.
\end{equation}
Since we can select $C$ to be as big as we want we can make this error as small as needed. This implies that then the main contribution to the integral~\eqref{eq:example_poisson_integral} comes from those `'heights" $y$ that are in the interval $\Kcal_C(k_n)$. In other words, the main contribution is concentrated around the heights $y$ for which $\mu(y) = k_n$. We thus refer to this as the concentration of heights result. More precisely, we prove the following.
